### 2022-01-27 analysis/paper/paper2.Rmd
Version with SNV performed before detrending for raman

---
title: "Semi-quantitative petrographic characterization of Bronze Age quartzite tools – a multivariate approach"
author:
  - Mattias Sjölander:
      email: mattias.sjolander@umu.se
      institute: [UmU]
      correspondence: true
institute:
  - UmU: Umeå University
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" 
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
mainfont: Times New Roman
bibliography: references.bib
csl: "../templates/journal-of-archaeological-and-anthropological-science.csl"
abstract: |
  Text of abstract
keywords: |
  Spectroscopy; Quartzite; Characterization; Lithics
highlights: |
  These are the highlights. 
---

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`

<!-- The following code chunk defines some general settings how code chunks should behave. -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)
```

<!-- The following code chunk defines prepares most of the data that will be needed and reads in the necessary libraries -->

```{r load-packages-and-data}
library(ChemoSpec)
library(data.table)
library(factoextra)
library(ggbiplot)
library(ggplot2)
library(ggh4x)
library(ggsci)
library(grid)
library(gridBase)
library(gridExtra)
library(missMDA)
library(patchwork)
library(prospectr)
library(tidyverse)
library(spectrolab)

#Import descriptive metadata
metadata.csv <-
  read.csv2("C:/Users/sjola/Documents/VERSION CONTROL/quantifying-quartzite/analysis/data/raw_data/metadata_20220119.csv", sep = ";", header = TRUE, na = c("", "NA", "NULL"))

#Import xrf data, set empty fields to NA
XRF.csv <-
  read.csv2("C:/Users/sjola/Documents/VERSION CONTROL/quantifying-quartzite/analysis/data/raw_data/XRF/XRF_quantitative.csv", sep = ";", dec = ".", header = TRUE, check.names = FALSE, na = c("","NA","NULL"))

#Import nir data, set empty fields to NA
nir.csv <-
  read.csv2("C:/Users/sjola/Documents/VERSION CONTROL/quantifying-quartzite/analysis/data/derived_data/NIR_master_averaged_20220119.csv", sep = ";", dec = ",", header = TRUE, check.names = FALSE, na = c("","NA","NULL",NULL))

#Import raman data, set empty fields to NA
raman.csv <-
  read.csv2("C:/Users/sjola/Documents/VERSION CONTROL/quantifying-quartzite/analysis/data/raw_data/RAMAN/raman_samples_data_non_treated_20220127.csv", sep = ";", dec = ",", header = TRUE, check.names = FALSE, na = c("","NA","NULL",NULL))

#Import sample list, set empty fields to NA
samples <-
  read.csv2("C:/Users/sjola/Documents/VERSION CONTROL/quantifying-quartzite/analysis/data/raw_data/bifacial_points.csv", sep = ";", dec = ".", header = TRUE, check.names = FALSE, na.strings=c("", "NA"))

#Filter xrf, nir and raman data to focus on points and preforms made from quartz/quartzite material 
Points.xrf <-
  XRF.csv %>%
    filter(material == "Quartz" | material == "Quartzite" | material == "Brecciated quartz" | material == "Brecciated quartzite" | material  == "Felsitic porphyric quartz",
           type == "Point" | type == "Point fragment" | type == "Preform")

Points.nir <-
  nir.csv %>%
    filter(material == "Quartz" | material == "Quartzite" | material == "Brecciated quartz" | material == "Brecciated quartzite" | material  == "Felsitic porphyric quartz",
           type == "Point" | type == "Point fragment" | type == "Preform")

#Fill the NA fields in the munsell_hue column and mark them as colourless samples
Points.nir[8][is.na(Points.nir[8])] <- "Colourless"

Points.xrf[20][is.na(Points.xrf[20])] <- "Colourless"

#prepare color/fill and symbols for score plots
pca.colors <- c("#E31CFF", "#718EFF", "#0080FF", "#D5DF9E", "#C6DF39", "#FFFFC1", "#FFFFA1", "#FFFF00", "#FFFF38", "#FFB468", "#000000", "#202020", "#404040", "#606060", "#808080", "#9F9F9F", "#DFDFDF", "#FFFFFF")
pca.shapes <- c(24,25,23,22,21)
pca.munsell_hue <- c("Colourless", "2.5PB", "10B", "2.5GY", "5GY", "2.5Y", "5Y", "7.5Y", "10Y", "5YR", "N1", "N2", "N3", "N4", "N5", "N6", "N8", "N9")

#If including references
#pca.colors <- c("#E31CFF", "#718EFF", "#0080FF", "#D5DF9E", "#C6DF39", "#FFFFC1", "#FFFFA1", "#FFFF00", "#FFFF38", "#DF5555", "#FFB468", "#000000", "#202020", "#404040", "#606060", "#808080", "#9F9F9F", "#BFBFBF", "#DFDFDF", "#FFFFFF")
#pca.munsell_hue <- c("Colourless", "2.5PB", "10B", "2.5GY", "5GY", "2.5Y", "5Y", "7.5Y", "10Y", "2.5YR", "5YR", "N1", "N2", "N3", "N4", "N5", "N6", "N7", "N8", "N9")

```

# Introduction

The last decades have seen an increase in the interest for non-flint lithics and raw material studies in European archaeology, providing much needed development in a previously neglected field [@blomme_etal12; @dalpra_pitblado16; @prieto_etal19; @prieto_etal20; @prieto_etal21; @ramacciotti_etal19]. In regions where local flint and obsidian sources are non-existent, such as Northern Fennoscandia, other materials have been the primary target for lithic production^[This is not to say that flint was not available or used in these regions] (e.g. quartz, quartzite, slate) [@baudou78; @rankama_etal06; @olausson_etal12]. With the lack of any natural geological source of flint in Northern Sweden both quartz and quartzite are instrumental for the inference of settlement and mobility within the region. While quartz and quartzite technology have been part of a number of studies over the years, there have been less development in the petrographic characterization and provenience studies of these same materials [@broadbent79; @callahan_etal92; @knutsson_etal16; @tallavaara_etal10]. One reason for the lack of methodological development in North Swedish research could be the complex geological background of the region and the abundance of moraine formations [@bargel03; @stephens_bergmanweihed20; @stroeven_etal16]. The degree to which geological materials have been mixed and transported demands considerable effort in order to reconstruct the events that have formed the modern day landscape. While tentaive interpretations of lithic material sourcing tends to describe strategies leaning towards direct procurement from the mountain zone [@forsberg85; -@forsberg12], both quartz and quartzite could potentially have been sourced locally from exposed veins in the bedrock, or as loose rocks and pebbles in the numerous moraine formations, as part of a strategy leaning more towards an embedded adaptation [@binford79; @rankama_etal06 pp.249-250]. Due to this ambiguous nature of availability of raw materials provenience studies are complex and fraught with uncertainty.

Prieto et al. [-@prieto_etal19 p.15; -@prieto_etal20 p.32] notes that the usage of the term *quartzite* in archaeology has been somewhat ambiguous, often used interchangeably between sedimentary orthoquartzite and metamorphic quartzite. It is difficult to provide a comprehensive definition of quartzite, owing much to its metamorphic nature, and different researchers have thus emphasized different features as being characteristic of the material [@howard05]. Visual characterization on a megascopic level is typically unreliable, and microscopic analysis is recommended for the identification of metamorphic structures, but Howard [-@howard05 p.708] also emphasises the need for field-based methodologies. In the case of archaeological research it is worth considering situations where destructive techniques simply are not applicable. Artefact collections (museum, private, etc.) represent a valuable resource and asset, with easy access to large datasets with a wide geographical distribution. While destructive methods may prohibit access to such material there are still non-destructive alternatives that can generate valuable information. Aside from being non-destructive spectroscopic instrumentation feature fast measurements, making it an attractive option for field-based screening purposes. While X-Ray Fluorescence (XRF) is the go-to spectroscopic technique for petrographic characterization [@shackley11a; @tykot03], due to its capability of providing compositional data, there are other instrumentations that may provide a deeper insight into the structural properties. The use of near-infrared spectroscopy (NIRS) goes almost as far back as the XRF [@hunt77], and there is an impressive body of work dedicated to the fingerprinting of geological materials [@linderholm_geladi14a; @linderholm_geladi14b; @sciuto19]. The same can be said for Raman spectroscopy, which has been applied extensively in the Cultural Heritage sector and in the analysis of artwork and manuscripts [@vandenabeele12].

\linebreak

## Aims

In the current study a dataset of quartz and quartzite artefacts from museum collections in Västerbotten, Sweden, have been analysed using the three above mentioned spectroscopic instrumentations (NIRS, Raman and XRF). The aim of the study is to evaluate their viability for fingerprinting quartz and quartzite material in the North Swedish region based on surface measurements. The study is a part of a larger project wherein models of settlement and mobility are being evaluated and developed in regards to raw material procurement and use. As flint is a non-local material to Northern Sweden it follows that any find of this material within the region most likely have been brought there through anthropogenic activity. In contrast, determining the degree to which cultural and/or environmental factors have influenced the distribution pattern of quartz and quartzite artefacts, however, is far more difficult. While there is a good understanding of the geological makeup of the Scandes [@stephens_bergmanweihed20], the quartzite itself remains largely uncharacterized, with few reference samples available for spectral comparison. The glacial impact on the landscape further compounds the issue as geological material has been transported and deposited by glacial movements, as well as subsequent postglacial processes. Although models for settlement and mobility can be used to test ideas of material distribution as influenced by cultural processes [e.g. @forsberg85], it is also necessary to attempt to estimate the degree of environmental impact on this pattern.

In order to engage in any meaningful analysis, however, it is first necessary to distinguish between different quartz and quartzite groups. Principal Component Analysis (PCA) has been used to model the spectral data in order to discern material groupings related to different spectral features and their contributions to the Principal Components (PC). The results form the basis of a discussion of the viability of using non-destructive spectroscopy in petrographic characterization of quartz and quartzite material, and thus future efforts of developing models for raw material sourcing in Northern Sweden.

\linebreak

<!--# Study region

\linebreak
 \linebreak

```{r geo-map, fig.width=15, fig.height=15, fig.fullwidth=TRUE, fig.cap="Geological map of Västerbotten County, Sweden, showing both the tectonic units of the mountains and the widely distributed moraine. Made with Lantmäteriet (CC0), Natural Earth, and Sveriges Geologiska Undersökning (SGU)"}
knitr::include_graphics('../figures/geo_map2.png', error = FALSE)
```

\linebreak

The effects of the last glacial on the landscape is visible in the widely distributed moraine formations and glaciofluvial deposits (fig. \@ref(fig:distribution-map)). In Västerbotten the Köli Nappe complex exhibits greensschist - amphibolite facies metamorphism, overlying the Seve Nappe complex where High Pressure (HP) and Ultra High Pressure (UHP) metamoprhism has been detected in a number of localities.

In 1925 the Geological Society arranged an exursion to Södra Storfjället in Västerbotten. In the description of the regional stratigraphy it is noted that the lower parts are made up of, among others, "*...dark schists and graphitic quartzite...*" [@geologiska25 p.147]. Apart from these quartzite with little impurities and Alum shale are also noted. 
-->
# Material
```{r select-point-data}
#Filter data to focus on points, point fragments and preforms made from quartz/quartzite materials
Points2 <-
  samples %>%
    filter(material == "Quartz" | material == "Quartzite" | material == "Brecciated quartz" | material == "Brecciated quartzite" | material  == "Felsitic porphyric quartz",
         type == "Point" | type == "Point fragment" | type == "Preform")

#Reorder type factor levels
Points2$type <-
  factor(Points2$type, levels=c("Preform", "Point fragment", "Point"))

#Reorder parish factor levels
Points2$parish <- 
  factor(Points2$parish, levels = c("Bjurholm","Bureå","Burträsk", "Bygdeå", "Degerfors", "Dorotea", "Lycksele", "Norsjö", "Skellefteå", "Sorsele", "Stensele", "Tärna", "Umeå", "Vännäs", "Vilhelmina", "Åsele"))
```

```{r sample-count, results='hide'}
#Filter the sample data to exclude reference samples
tot_samples <- 
  samples %>% 
    filter(type != "Reference")

#Filter sample data to focus on site Lycksele 301
lycksele <-
  samples %>% 
    filter(site_id == "Lycksele 301")

#Then filter to focus on quartz/quartzite material that are points and preforms 
lycksele_301 <-
  lycksele %>% 
    filter(material == "Quartz" | material == "Quartzite" | material == "Brecciated quartz" | material == "Brecciated quartzite" | 
             material  == "Felsitic porphyric quartz", type == "Point" | type == "Point fragment" | type == "Preform")

#Create vector of unique samples from site Lycksele 301
lycksele_301 <-
unique(lycksele_301$sample_id)

#Create a vector containing all unique sample ID, in order to count number of total samples (incl. types other than points/preforms or materials) 
tot_uniq <- 
  unique(tot_samples$sample_id)

#Create a vector in order count number of unique rows in sample ID column for the number of points/preforms included in analysis
sample_uniq <- 
  unique(Points2$sample_id)
```

\linebreak
 \linebreak

The current study is based on spectroscopic surface measurements collected on quartz and quartzite points recovered from sites within the County of Västerbotten, Sweden. A dataset of `r length(sample_uniq)` points, point fragments and preforms made from quartz and quartzite material stored at the collections at Västerbotten (www.vbm.se) and Skellefteå (www.skellefteamuseum.se) museums were analysed using NIRS, raman scattering and XRF. The material features a mainly black, grey or white hue, as well as a subset of semi-translucent material with a beige tint to almost fully translucent. The colour of the material has been classified both according to the Munsell system, as well as sorted into general categories of hue for visualization purposes: *dark* (~ N1-N4), *light* (~ N5-N8), *white* (~ N9), *colourless* (lack of colour, opaque to translucent). A number of these points have, according to prior classifications, been worked using bifacial technology. While there currently is no local chronology available for this artefact type in the North Swedish region^[For comparison, Mjæreum -@mjaerum12 has put forth a chronology for bifacial points in southwestern Norway], there have been attempts [@baudou92 p.99; @forsberg85 p.5; @forsberg89]. Previous research suggests a period of use that falls between ca. 3 800 - 2 000 BP [@kaeck09 p.60]. Similarly, studies of raw material use at sites dating to this period seems to indicate an increase in the use of quartzite [@baudou78].

A large portion of the collections stored at North Swedish museums were recovered from excavations and surveys performed in conjunction with the river regulations during the 1940s – 1980s [@biornstad06; @janson_hvarfner60], as well as subsequent follow-up survey projects [e.g. @andersson15]. This has likely had a contributing factor to the distribution pattern of the artefacts observed in fig. \@ref(fig:distribution-map) as the known sites with finds of points and preforms are concentrated around the parishes of Dorotea, Lycksele^[In this case the majority of the samples from Lycksele are from the rock painting site of Lycksele 301], Vilhelmina and Tärnaby. This can be seen more clearly in fig. \@ref(fig:distribution-by-parish) with the four parishes accounting for almost ¾ of the total no. of samples.

Västerbotten museum have come far in their digitization effort, making large portions of their collections searchable online. Based on a query of all points and preforms labelled as *quartz*, *brecciated quartz* or *quartzite* a list was generated from their database. Sampling of this list of artefacts was not random. Sites located in areas with few finds were included first in order to increase the geographical distribution, with sampling of areas with higher site density making up the remaining bulk of the dataset. None of these were from the northernmost Skellefte river, however, and thus a selection of 5 artefacts each from three well-known sites stored at the collections of Skellefte museum were included. This resulted in a total dataset of `r length(tot_uniq)` artefacts, and after documentation and looking up missing information, `r length(sample_uniq)` of these were determined to be either points or preforms made from quartz or quartzite material.

\linebreak

```{r distribution-map, fig.width=15, fig.height=15, fig.fullwidth=TRUE, fig.cap="Geographical distribution of sampled sites within Västerbotten County, Sweden. The distribution of sampled sites is likely affected by the excavation and survey history of the region. White: Sampled sites; Black: Parish centroid for measured samples lacking site coordinates. Made with Lantmäteriet (CC0), Natural Earth, and Riksantikvarieämbetet."}
knitr::include_graphics('../figures/Distribution.jpeg', error = FALSE)
```

\linebreak

```{r distribution-by-parish, fig.width=12, fig.height=10, fig.fullwidth=TRUE, fig.cap= paste("Distribution of the", length(sample_uniq), "points and preforms across the parishes in Västerbotten, Sweden. Note that while Lycksele feature a large number of samples", length(lycksele_301), "of these are from the site of Lycksele 301")}

#Stacked bar chart with counts of samples per parish and type
ggplot(Points2, aes(x= fct_rev(parish), fill = type)) +
  geom_bar() +
  theme_minimal() +
  labs(y = "Count",
       fill = "Type") +
  theme(
    legend.position = "bottom",
    legend.title = element_text(size = 16, face = "bold"),
    legend.text = element_text(face = "bold"),
    axis.title.y=element_blank(),
    axis.text.y = element_text(size = 10, face = "bold", colour = "black"),
    axis.title.x = element_text(size = 12, face = "bold", colour = "black")) +
  scale_fill_jco() +
  coord_flip() +
  geom_text(stat ='count', aes(group = parish, label = ..count..), hjust = -0.5)
```

\linebreak

While parts of the collections at Västerbotten museum have been petrographically analysed by a geologist [@biornstad68; @ahman67], at this point it is difficult to say with any certainty how many (if any) of the artefacts in the current dataset were among these. It is fair to assume that the vast majority have been classified visually based on megascopic features according to a “standard” archaeological nomenclature. In the dataset the four main types of registered materials are: quartz, quartzite, brecciated quartz and brecciated quartzite^[Original Swedish terms are *brecciekvarts*/*brecciekvartsit*]. Those that had no prior material classification registered were sorted into either the quartz or quartzite category. The reason for including quartz in the study is in order to test whether it is possible to separate quartz from quartzite using these methods, as well as a way to account for potential mistakes in the original classification.

Another reason for keeping a quartz category is the group referred to as *brecciated quartz*. In a petrographic study of the artefacts recovered along the Ume river valley Åhman [-@ahman67 p.8] briefly touches upon the definition of *brecciated quartz*, also referring to it as a *mylonite quartz*. Åhman describes the material as having formed from a quartz vein which has fractured due to movements in the mountain massives and recrystallized via pressure. While the translucent material may give the impression of an unusually pure quartzite, Åhman states that the thin-section analysis reveals the fragmented configuration of quartz. *Brecciated quartz* is a somewhat common term used in North Swedish archaeology, and has typically been used to identify a certain translucent type of fine-grained, quartz-rich material [e.g. @holm91 p.24].

\linebreak

# Methods

\linebreak
 \linebreak

All measurements were performed at the Environmental Archaeology Lab at Umeå University^[www.umu.se/forskning/infrastruktur/mal/]. The analysis was conducted in a dark room in order to minimize light pollution, with the main light sources in the room being the computer screens. Each artefact was sampled twice, on different sides, with additional measurements in cases where sampling was difficult. The only exception to this was the XRF, where due to time constraints one measurement was collected as a rule apart from a few exceptions. This lowers the statistical reliability of the XRF analysis and inferences of the results should be evaluated with this in mind. The size of the dataset alleviates some of these issues, as for the larger sites there are multiple objects that may be of the same material, however, for the smaller sites there still remains some uncertainties. For materials that featured high heterogeneity the sample point was focused, as much as possible, on the surrounding matrix.

Pre-processing of the spectra is different for each instrumentation and based on relevant literature. After pre-processing the data was modelled using PCA, which is an unsupervised exploratory technique for reducing the dimensionality of the dataset [@geladi_linderholm20]. PCA modelling was performed using the base R *prcomp()* function [@Rteam17; @rcoreteam21].

## Near Infrared
NIRS measures the absorbance of light within the near infrared region (780 nm – 2 500 nm) of the electromagnetic spectrum. Based on the theory of molecular vibration and excitation [@hunt77], by exposing a geological material to electromagnetic radiation it is possible to generate overtone vibrations which can be related to certain molecular groups. The most common group capable of generating these overtones is the OH group, which in quartz-based materials typically relates to molecular water [@hunt77 p.508].

NIR analysis was conducted with the Analytical Spectral Device (ASD) LabSpec 4, using a contact probe. The ASD features a detection range of 350 nm – 2 500 nm, with spectral sampling (resp. spectral resolution) of 1.4 nm (resp. 3 nm) in the visible and near infrared range and 1.1 nm (resp. 10 nm) in the short‐wave infrared range. Each object was measured against a white reference background and the collected measurements were averaged for each artefact. The visible spectrum (~ 400 nm - 750 nm) and the NIR range (~ 1 000 nm – 2 500 nm) were analysed separately, with the former focused on colour analysis and the latter the chemical structure of the material. The data was mean-centred before PCA modelling.

\linebreak

## Raman
Raman spectroscopy shares some similarities with NIRS in that they both can be used to infer molecular structure based on their vibrational properties [@nafie01]. However, where NIRS measures light absorbance and generate overtones of molecular vibration, Raman spectroscopy is based on the *Raman scattering* effect. By exposing a molecule to radiation with a monochromatic laser it is possible to induce a dipole moment in the molecule which generates the Raman effect. The subsequent light that is scattered contains both Rayleigh scatter and Raman scatter, the former at a frequency of incident radiation and the latter at a shifted frequency [@dubessy12]. The spectrometer will measure the Raman scatter and generate a spectrum where the spectral line has shifted compared to the incident Rayleigh frequency. As the structure of a molecule will inform its vibrational mode, the point at which these shifts occur within the infrared region varies depending on the material [@smith_carabatos-nedelec01].

The Raman instrumentation used was a portable i-Raman EX featuring a 1064 nm excitation laser. The spectral coverage of the instrumentation is 175 cm^-1^ – 2 500 cm^-1^ ^[Raman shifts are typically given in wavenumber (cm^-1^), instead of spectral wavelength (nm)] with a resolution of 9.5 cm^-1^. Access to the spectrometer was made possible by the department of Forest Biomaterials and Technology at the Swedish University of Agricultural Sciences (SLU) in Umeå^[www.slu.se/en/departments/forest-biomaterials-technology/].

After averaging the spectra for each artefact the results were baseline corrected in order to detrend the data. This was achieved using modified polynomial fitting in R [@Rteam17] with the ChemoSpec package [@hanson16]. The spectra were then mean-centred and normalized using Standard Normal Variate [@shaver01 pp.284-285] before PCA.

\linebreak

## X-Ray Fluourescence
XRF spectroscopy, and its capabilities for quantifying chemical elements in different materials, has made it a well-established method for petrographic analysis in archaeology [@hall60; @shackley11a]. Exposing a material to X-ray radiation causes electrons to dislodge and move between the electron shells of the atoms, which in turn emits energy (i.e. *fluourescence*). As the energy differences between electron shells for the various elements is known it is possible to both identify and quantify the abundance of an element [@shackley11b p.16], which is useful for provenience studies where the chemical character of an unknown and a reference sample can tested statistically [@tykot03].

XRF measurements were collected using a Thermo Scientific Niton XL5 Analyzer, connected to a portable test stand. The reference calibration used for element quantification was set to *mining* mode. Sampling with this setting takes 2 minutes in order for the instrument to cycle through all of the filters (main, low, high, light). Due to time constraints and the number of objects in the current dataset it was only possible to collect one measurement on each object, apart from some exceptions.

The XRF data was log-transformed, mean-centered and scaled to unit variance before PCA modelling. As the resulting dataset also features values below the Limit of Detection (LOD), these were treated as NULL (or "missing values"), during PCA. Since missing values are not permitted during PCA these had to be imputed before the analysis, which was performed using the *missMDA* R package [@josse_husson16] developed for datasets with incomplete data. Two methods for imputation can be implemented, iterative or regularized, and since the former suffers when there is a large number of missing values the latter was used.

\pagebreak

# Results

\linebreak
 \linebreak

## NIR - visual range
Examining the average spectra of the different colour groups (dark, light, white and colourless), as expected, the darker material feature higher absorbance of light than the other groups. No other distinctive features can be seen in the visual band, that could potentially be used to distinguish them. It is, however, apparent that while the more extreme hues of dark and white material can be isolated more clearly based on their spectra, there also exists a wider greyscale of material that does not conform as easily to these broader groups.

The cumulative explained variance for the first three components in the PCA is ~ 99.9%, with the accompanying loadings in fig. \@ref(fig:vis-nir-loadings). Most of the contribution to the first Principal Component (PC) can be found within the red part of the spectrum (~ 600 nm - 750 nm). This can be seen in the score plot where darker and lighter materials are separated along PC 1, as darker samples would feature a higher absorbance within the red part of the spectrum (fig. \@ref(fig:vis-nir-pca)). In contrast, most of the contribution to PC 2 occurs within the blue-violet part of the spectrum, which does not seem to alter the structure of the dataset in any significant way. PC 3 feature high contributions within both the blue-violet and red parts of the spectrum, which also seems to draw the light and dark samples that feature a blue-gray tint towards the top of the score plot.

While the overall structure of the model seems to group colourless and white materials towards the bottom left, with dark material to the right, there are a number of samples that break this pattern. This suggests that the perceived colour of a material does not necessarily match what is detected within the visible spectrum of light. It is also difficult to separate out the greyscale of colours that occur inbetween the "black and white" groups, where both light gray and semi-translucent samples end up. All in all, the results echo the complexity and unreliability that exists within the topic of petrographic colour.

\linebreak

```{r vis-nir-loadings, fig.width=20, fig.height=15, fig.cap="Loading lines for PC 1 - 3. The two first components mainly describes features with a linear increase in absorbance across the visual spectrum, but the 3rd component demonstrates more variation, with lower absorbance within the blue-yellow region. "}

#perform PCA with SNV normalization and mean-center
visnir.pca <-
  prcomp(Points.nir[,c(78:428)], center = TRUE, scale = FALSE)

#Filter NIR data to focus on the material with a dark hue 
#and select the NIR range 1 000 - 2 500 nm
#vis.dark <- Points.nir %>% 
#  filter(hue == "Dark") %>% 
#  select(c(1:28),c(79:429))

#spec.vd <- 
#      as_spectra(vis.dark, name_idx = 1, meta_idxs = 2:28)

#vis.light <- Points.nir %>% 
#  filter(hue == "Light") %>% 
#  select(c(1:28),c(79:429))

#spec.vl <- 
#  as_spectra(vis.light, name_idx = 1, meta_idxs = 2:28)

#vis.colourless <- Points.nir %>% 
#  filter(hue == "Colourless") %>% 
#  select(c(1:28),c(79:429))

#spec.vt <- 
#  as_spectra(vis.colourless, name_idx = 1, meta_idxs = 2:28)

#vis.white <- Points.nir %>% 
#  filter(hue == "White") %>% 
#  select(c(1:28),c(79:429))

#spec.vw <- 
#  as_spectra(vis.white, name_idx = 1, meta_idxs = 2:28)

#VISNIR LOADING LINES
#For some reason invisible() has to be used to prevent console output showing up in knitted document
par(mfrow=c(3,1))
invisible(sapply(1:3, function(comp){
  plot(400:750, visnir.pca$rotation[,comp], type='l', lwd=2,
      main=paste("PC", comp), xlab="Wavelength (nm)", ylab="", cex.lab = 1.2, xaxp=c(400,750,5))}))

#par(mfrow=c(4,1))

#plot(mean(spec.vd), lwd = 1.2, col = "black", ylab = "Absorbance", cex.lab = 1.5, cex.main = 1.5, xlab = "Wavelength (nm)", main = "Average - Dark"); vps1 <- do.call(vpStack, baseViewports())

#plot(mean(spec.vl), lwd = 1.2, col = "red", ylab = "Absorbance", cex.lab = 1.5, cex.main = 1.5, xlab = "Wavelength (nm)", main = "Average - Light")

#plot(mean(spec.vt), lwd = 1.2, col = "purple", ylab = "Absorbance", cex.lab = 1.5, cex.main = 1.5, xlab = "Wavelength (nm)", main = "Average - Colourless")

#plot(mean(spec.vw), lwd = 1.2, col = "blue", ylab = "Absorbance", cex.lab = 1.5, cex.main = 1.5, xlab = "Wavelength (nm)", main = "Average - White"); vps4 <- do.call(vpStack, baseViewports())
```

\linebreak

```{r vis-nir-pca, fig.width=20, fig.height=12, fig.cap="PCA score plots for the 400 nm – 750 nm band of the NIR data. Top: PC 1 and PC 2, bottom: PC 1 and PC 3. PC 1 seems to group dark-light material on opposite sides of the plots while PC 2 does little to change the strucure. PC 3 seems to position material with a blue-gray tint towards the top of the score plot."}

#prepare a basic score plot with fviz_pca_ind using PC1 and PC2
basic_plot1 <-
  fviz_pca_ind(visnir.pca, axes = c(1,2), label="none")
basic_plot2 <-
  fviz_pca_ind(visnir.pca, axes = c(1,3), label="none")

#prepare labels for PCs
vis.pc1var <- round(summary(visnir.pca)$importance[2,1]*100, digits=2)
vis.pc2var <- round(summary(visnir.pca)$importance[2,2]*100, digits=2)
vis.pc3var <- round(summary(visnir.pca)$importance[2,3]*100, digits=2)
vis.pc1lab <- paste0("PC1 (",as.character(vis.pc1var),"%)")
vis.pc2lab <- paste0("PC2 (",as.character(vis.pc2var),"%)")
vis.pc3lab <- paste0("PC3 (",as.character(vis.pc3var),"%)")

#VISNIR LOADING LINES
#par(mfrow=c(3,1))
#sapply(1:3, function(comp){
#  plot(400:750, visnir.pca$rotation[,comp], type='l', lwd=2,
#      main=paste("PC", comp), xlab="Wavelength (nm)", ylab="", cex.lab = 1.2, xaxp=c(400,750,6))})

#bind the basic fviz plot for PC 1 and 2, and use as basis for a more customizeable plot in ggpplot
visnir.p1 <-
  ggplot(cbind(basic_plot1$data, Points.nir[, c(8,10)]),
             aes(x=x, y=y, 
                 shape = material, 
                 fill = munsell_hue)) +
  geom_point(size=3) +
  theme_bw() +
  ggtitle("Near infrared 400 - 750 nm") +
  labs(x = vis.pc1lab,
       y = vis.pc2lab) +
  scale_shape_manual(name = "Material",
                     values=pca.shapes) +
  scale_fill_manual(name = "Munsell hue",
                    breaks = c("Colourless", "2.5PB", "10B", "2.5GY", "5GY", "2.5Y", "5Y", "7.5Y", "10Y", "5YR", "N1", "N2", "N3", "N4", "N5", "N6", "N8", "N9"),
                    values=pca.colors) +
  guides(shape = guide_legend(override.aes = list(fill = "black"), 
                              ncol=2, 
                              title.position="top", 
                              title.hjust = 0.5),
         fill = guide_legend(override.aes = list(shape = 22,
                                                 fill = pca.colors,
                                                 color = "black",
                                                 size=3),
                             title.position="top", 
                             title.hjust = 0.5)) +
  theme(legend.position = "none")

visnir.p2 <-
  ggplot(cbind(basic_plot2$data, Points.nir[, c(8,10)]),
             aes(x=x, y=y, 
                 shape = material, 
                 fill = munsell_hue)) +
  geom_point(size=3) +
  theme_bw() +
  ggtitle("Near infrared 400 - 750 nm") +
  labs(x = vis.pc1lab,
       y = vis.pc3lab) +
  scale_shape_manual(name = "Material",
                     values=pca.shapes) +
  scale_fill_manual(name = "Munsell hue",
                    breaks = c("Colourless", "2.5PB", "10B", "2.5GY", "5GY", "2.5Y", "5Y", "7.5Y", "10Y", "5YR", "N1", "N2", "N3", "N4", "N5", "N6", "N8", "N9"),
                    values=pca.colors) +
  guides(shape = guide_legend(override.aes = list(fill = "black"), 
                              ncol=2, 
                              title.position="top", 
                              title.hjust = 0.5),
         fill = guide_legend(override.aes = list(shape = 22,
                                                 fill = pca.colors,
                                                 color = "black",
                                                 size=3),
                             title.position="top", 
                             title.hjust = 0.5)) +
  theme(legend.position = "none")

comb1 <-
  visnir.p1 + visnir.p2 & theme(plot.title = element_text(size = 12, face = "bold", colour = "black", vjust = - 10, hjust = 0.02),
                              axis.title.x = element_text(size = 12, face = "bold", colour = "black"),
                              axis.title.y = element_text(size = 12, face = "bold", colour = "black"),
                              legend.title = element_text(size = 12, face = "bold", colour = "black"),
                              legend.background = element_rect(linetype = "solid", color = "black"),
                              legend.position = "bottom")

comb1 +
  plot_layout(nrow = 2, guides = "collect")
```

\linebreak

## NIR - near infrared range
In the case of the 1 000 nm - 2 500 nm band of the NIR data, similar features can be observed in the average spectra for all four colour groups, with darker materials exhibiting higher absorbance and the light group the most distinct peaks. The presence of molecular water can be inferred based on the peaks at 1 400 nm and 1 900 nm  [@hunt77; @sciuto_etal19] (fig. \@ref(fig:nir-spectra)). Peaks also occur at 2 200 nm and 2 350 nm, where the former has been related to combination vibrations of Al-OH and the latter falling within one of the stronger CO^3^ bands [@clark_etal90]. Inspection of individual spectra for the light group, however, also reveals a 5th peak for some samples at 2 450 nm. Comparisons to spectra collected on muscovite [@clark_etal90 p.12 675] shows similar peaks located at 2 200 nm, 2 350 nm, and 2 450 nm. It is therefore possible that the 2 350 nm band in some cases corresponds to CO^3^ vibrations and in other samples Al-OH, possibly inidicating the presence of muscovite.

The PCA shows a strong trend along PC 1 that seems to correlate with the colour of the material(fig. \@ref(fig:nir-pca)). Darker quartzite cluster to the right with white and colourless samples grouping to the opposite side. While the white samples seem to gravitate to the bottom left of the model the samples that are somewhat translucent, such as the brecciated quartz, seem to cluster towards the top-left. At the top of the score plot 3 black samples separate themselves clearly from the rest of the data points. These three samples come from different sites and exhibit no peaks in their spectra, making them an interesting set of outliers.

\linebreak

```{r nir-spectra, fig.width=20, fig.height=15, fig.cap="Average NIR spectra by hue. Darker samples feature higher overall absorbance, as well as less pronounced peaks, when compared to lighter/colourless samples."}

#Filter NIR data to focus on the material with a dark hue 
#and select the NIR range 1 000 - 2 500 nm
dark <- Points.nir %>% 
  filter(hue == "Dark") %>% 
  select(c(1:28),c(680:2179))

#create a spectra object for the dark material, and then repeat for light, white and colourless groups
spec.d <- 
  as_spectra(dark, name_idx = 1, meta_idxs = 2:28)

light <- Points.nir %>% 
  filter(hue == "Light") %>% 
  select(c(1:28),c(680:2179))

spec.l <- 
  as_spectra(light, name_idx = 1, meta_idxs = 2:28)

colourless <- Points.nir %>% 
  filter(hue == "Colourless") %>% 
  select(c(1:28),c(680:2179))

spec.t <- 
  as_spectra(colourless, name_idx = 1, meta_idxs = 2:28)

white <- Points.nir %>% 
  filter(hue == "White") %>% 
  select(c(1:28),c(680:2179))

spec.w <- 
  as_spectra(white, name_idx = 1, meta_idxs = 2:28)

#Prepare a layout with 1 column and four rows
par(mfrow=c(4,1))

#plot the mean spectra for each group
plot(mean(spec.d), lwd = 1.2, col = "black", ylab = "Absorbance", cex.lab = 1.5, cex.main = 1.5, xaxp=c(1000,2500,6),
     xlab = "Wavelength (nm)", main = "Average - Dark"); vps1 <- do.call(vpStack, baseViewports())

plot(mean(spec.l), lwd = 1.2, col = "red", ylab = "Absorbance", cex.lab = 1.5, cex.main = 1.5, xaxp=c(1000,2500,6),
     xlab = "Wavelength (nm)", main = "Average - Light")

plot(mean(spec.t), lwd = 1.2, col = "purple", ylab = "Absorbance", cex.lab = 1.5, cex.main = 1.5, xaxp=c(1000,2500,6),
     xlab = "Wavelength (nm)", main = "Average - Colourless")

plot(mean(spec.w), lwd = 1.2, col = "blue", ylab = "Absorbance", cex.lab = 1.5, cex.main = 1.5, xaxp=c(1000,2500,6),
     xlab = "Wavelength (nm)", main = "Average - White"); vps4 <- do.call(vpStack, baseViewports())

pushViewport(vps1)
X <- convertX(unit(1413,"native"), "npc")
X2 <- convertX(unit(1935,"native"), "npc")
X3 <- convertX(unit(2215,"native"), "npc")
X4 <- convertX(unit(2350,"native"), "npc")
#For some reason grid.text for peak labels does not always align with the x value for the dashed line, so need to add a specific value for the text
x1 <- convertX(unit(1436,"native"), "npc")
x2 <- convertX(unit(1936,"native"), "npc")
x3 <- convertX(unit(2212,"native"), "npc")
x4 <- convertX(unit(2335,"native"), "npc")
popViewport(3)

grid.move.to(x = X, y = unit(1, "npc"), vp = vps1)
grid.line.to(x = X, y = unit(0, "npc"), vp = vps4, 
             gp = gpar(col = "gray", lty = 2, lwd = 2))
grid.text("OH", x = x1, y = unit(0.87, "npc"), just="center", gp=gpar(font=2))

grid.move.to(x = X2, y = unit(1, "npc"), vp = vps1)
grid.line.to(x = X2, y = unit(0, "npc"), vp = vps4, 
             gp = gpar(col = "gray", lty = 2, lwd = 2))
grid.text("OH", x = x2, y = unit(0.93, "npc"), just="center", gp=gpar(font=2))

grid.move.to(x = X3, y = unit(1, "npc"), vp = vps1)
grid.line.to(x = X3, y = unit(0, "npc"), vp = vps4, 
             gp = gpar(col = "gray", lty = 2, lwd = 2))
grid.text("AlOH", x = x3, y = unit(0.95, "npc"), just="center", gp=gpar(font=2))

grid.move.to(x = X4, y = unit(1, "npc"), vp = vps1)
grid.line.to(x = X4, y = unit(0, "npc"), vp = vps4, 
             gp = gpar(col = "gray", lty = 2, lwd = 2))
grid.text("CO3", x = x4, y = unit(0.95, "npc"), just="center", gp=gpar(font=2))
```

\linebreak

```{r nir-pca, fig.width=20, fig.height=8, fig.cap= "PCA score plot of 1st and 2nd components for the 1 000 nm – 2 500 nm region of the spectrum. Quartz-based material with a darker surface group to the right in the model, while those with white surface group in the bottom left. A trend of translucent material and opaque material with a beige tint grouping in the top left can also be seen. Three black samples at the top, from different sites, all feature no distinguishing peaks in their spectra thus separating them from the other samples in the dataset."}

#perform PCA with SNV normalization and mean-center
nir.pca <-
  prcomp(Points.nir[,c(680:2179)], center = TRUE, scale = FALSE)

#prepare a basic score plot with fviz_pca_ind using PC1 and PC2
basic_plot1 <-
  fviz_pca_ind(nir.pca, axes = c(1,2), label="none")

#prepare labels for PCs
nir.pc1var <- round(summary(nir.pca)$importance[2,1]*100, digits=2)
nir.pc2var <- round(summary(nir.pca)$importance[2,2]*100, digits=2)
nir.pc1lab <- paste0("PC1 (",as.character(nir.pc1var),"%)")
nir.pc2lab <- paste0("PC2 (",as.character(nir.pc2var),"%)")

#NIR LOADING LINES
#par(mfrow=c(3,1))
#sapply(1:3, function(comp){
#  plot(1001:2500, nir.pca$rotation[,comp], type='l', lwd=2,
#      main=paste("PC", comp), xlab="Wavelength (nm)", ylab="", cex.lab = 1.2, xaxp=c(1001,2500,6))})

#bind the basic fviz plot for PC 1 and 2, and use as basis for a more customizeable plot in ggpplot
ggplot(cbind(basic_plot1$data, Points.nir[, c(8,10)]),
             aes(x=x, y=y, 
                 shape = material, 
                 fill = munsell_hue)) +
  geom_point(size=3) +
  theme_bw() +
  ggtitle("Near infrared 1 000 - 2 500 nm") +
  labs(x = nir.pc1lab,
       y = nir.pc2lab) +
  scale_shape_manual(name = "Material",
                     values=pca.shapes) +
  scale_fill_manual(name = "Munsell hue",
                    breaks = c("Colourless", "2.5PB", "10B", "2.5GY", "5GY", "2.5Y", "5Y", "7.5Y", "10Y", "5YR", "N1", "N2", "N3", "N4", "N5", "N6", "N8", "N9"),
                    values=pca.colors) +
  guides(shape = guide_legend(override.aes = list(fill = "black"), 
                              ncol=2, 
                              title.position="top", 
                              title.hjust = 0.5),
         fill = guide_legend(override.aes = list(shape = 22,
                                                 fill = pca.colors,
                                                 color = "black",
                                                 size=3),
                             title.position="top", title.hjust = 0.5))+
  theme(plot.title = element_text(size = 12, face = "bold", colour = "black", vjust = - 10, hjust = 0.02),
        axis.title.x = element_text(size = 12, face = "bold", colour = "black"),
        axis.title.y = element_text(size = 12, face = "bold", colour = "black"),
        legend.title = element_text(size = 12, face = "bold", colour = "black"),
        legend.background = element_rect(linetype = "solid", color = "black"),
        legend.position = "bottom")
```

\linebreak

## Raman
There are two distinctive features present in the Raman data (fig. \@ref(fig:raman-spectra)). The first are three peaks in the 100 cm^-1^ - 500 cm^-1^ band that are typical for SiO~2~ and quartz-based materials, with the strongest peak occuring at 460 cm^-1^ [@gillet_etal90]. These peaks are present for all samples in the dataset. The second feature shows up as two peaks occurring at around 1 287 cm^-1^ and 1 598 cm^-1^, and dominate the spectra in terms of intensity for the darker samples. These peak positions fall within what is referred to as the *D* and *G* band respectively, and have been observed in spectroscopic studies of graphite [@cesare_maineri99]. The D band occurs as a result of disordered carbon (C) structures whilst the G band is indicative of sp^2^ carbon networks and caused by stretching in the carbon bonds, which is common in graphite-based materials [@cohen-ofri_etal06; @vollebregt_etal12; @lee_etal21]. Whereas the 2^nd^ peak falls within the typical position of the G band, the position of the D band is commonly identified as being in the 1 330 - 1 350 cm^-1^ region. It has, however, been documented that the dispersive behavior of the D band can cause a frequency change as a result of the energy of the incident laser [@pimenta_etal07 p.1280]. This could thus explain the discrepancy between the observed peak in the dataset and the literature.

A third feature, present in a small subset of the samples, appears as two peaks spaced closely together at ~ 695 cm^-1^ and 795 cm^-1^. Comparing these to reference samples from the mineral and rock collections kept at the Department of Ecology and Environmental Science at Umeå University, the same peaks are present in measurements collected on, among others, albite, apatite, diabase and shale. Studies of clay minerals [@kloprogge17] have identified a number of band regions in the lower wavenumbers related to Si, Al and OH vibrations, including the 700 - 800 cm^-1^ band. Considering the petrogenesis of quartzite and the likelihood that clay minerals are present in the material it is possible that this is what generates the feature in the current dataset.

The resulting PCA model (fig. \@ref(fig:raman-pca)) shows a clear trend along PC 1 (81.5 %) related to the quartz and D - G peaks. PC 2 and PC 3, however, mainly describes the remaining noise left in the spectra after detrending and preprocessing the data. Interestingly, the loading for PC 4 does include the Al-O-Si peaks around the 700 cm^-1^ - 800 cm^-1^ band, which helps in separating out those samples where this feature is present. The measurements with mainly quartz peaks gravitates towards the negative side of PC 1, and those where the graphite peaks are dominant towards the opposite side. It is also clear that the darker material feature more pronounced graphite peaks, whereas the lighter and more colourless objects mainly feature quartz peaks. As in the case of the NIR data, however, there are exceptions that break this trend, as well as a “muddled” middle group with both well-pronounced graphite and quartz peaks. When PC 4 is added the samples with quartz and Al-O-Si peaks gravitate towards the bottom left, separating themselves from the "quarts" group.

\linebreak

```{r raman-prep-data}
#aggregate observations by group(sample) and calculate average of wavelength measurements
raman.averaged <- 
  aggregate(raman.csv[, 5:469], list(sample_id = raman.csv$sample_id), mean)

#perform Standard Normal Variate on raman data
raman.averaged[,2:466] <- 
  standardNormalVariate(raman.averaged[,2:466])

#merge data with metadata
#raman_merged <- 
#  as.data.frame(merge(metadata, raman_averaged, by='sample_id'))

#select sample_id and data columns and transpose in prep for baseline de-trend
raman.transposed <- 
  raman.averaged %>% 
    select(sample_id, `92.88`:`2503.59`)

raman.transposed <- 
  t(raman.transposed)

#Write csv with transposed raman data for baseline de-trending in ChemoSpec package
#ChemoSpec requires data to be read from file
write.table(raman.transposed, file = "C:/Users/sjola/Documents/VERSION CONTROL/quantifying-quartzite/analysis/data/derived_data/raman_transposed.csv", col.names = FALSE, sep = ";", dec = )

#prepare vector with sample_id for reading raman data into spectra object
sample_id <- 
  raman.averaged$sample_id

#read transposed raman data into spectra object
raman <-matrix2SpectraObject(
  gr.crit = sample_id,
  gr.cols = c("auto"),
  freq.unit ="Wavelength cm-1",
  int.unit ="Intensity",
  descrip ="Bifacial points measurements",
  in.file = "C:/Users/sjola/Documents/VERSION CONTROL/quantifying-quartzite/analysis/data/derived_data/raman_transposed.csv",
  sep = ";",
  dec = ".",
  chk = TRUE,
  out.file = "C:/Users/sjola/Documents/VERSION CONTROL/quantifying-quartzite/analysis/data/derived_data/raman_spec_object")

#call on relevant baseline method (modified polynomial fitting) from baseline package and 
#return corrected spectra to spectra object
raman.baseline <- 
  baselineSpectra(raman,
                  int = FALSE,
                  method = "modpolyfit",
                  retC = TRUE)

#spectra are stored in "Spectra" object, not a data frame
#need to extract samples and frequencies from that object 
raman.tmp <- 
  cbind(raman.baseline$freq, t(raman.baseline$data))
#transpose data to get frequencies as columns
raman.tmp <- 
  as.data.frame(t(raman.tmp))
#Add sample_id column and remove the initial "X" character introduced to sample names by R
raman.tmp$sample_id <- 
  c("sample_id", sub("^X*", "", raman.baseline$names))

#Make first row headers and remove the first row with the names
names(raman.tmp) <-
  raman.tmp[1,]
raman.tmp <- 
  raman.tmp[-1,]

#merge raman de-trended data with metadata
raman.baseline.merged <- 
  as.data.frame(merge(metadata.csv, raman.tmp, by='sample_id'))

#select and filter raman data to focus on relevant material and artefact types
Points.raman <-
  raman.baseline.merged %>%
  filter(material == "Quartz" | material == "Quartzite" | material == "Brecciated quartz" | material == "Brecciated quartzite" | material  == "Felsitic porphyric quartz",
         type == "Point" | type == "Point fragment" | type == "Preform")

#fill the NA fields in the munsell hue column to mark them as colourless/translucent material
Points.raman[8][is.na(Points.raman[8])] <- "Colourless"
```

```{r raman-spectra, fig.width=20, fig.height=15, fullwidth = TRUE, fig.cap="Average Raman spectra grouped by their hue. The *quartz peak* at 460 cm^-1^ is present in all four groups, but the *graphite peaks* at 1 287  cm^-1^ (D band) and 1 598 cm^-1^ (G band) mainly appear in the dark group"}

r.dark <- Points.raman %>% 
  filter(hue == "Dark")

spec.rd <- 
  as_spectra(r.dark, name_idx = 1, meta_idxs = 2:29)

r.light <- Points.raman %>% 
  filter(hue == "Light")

spec.rl <- 
  as_spectra(r.light, name_idx = 1, meta_idxs = 2:29)

r.colourless <- Points.raman %>% 
  filter(hue == "Colourless")

spec.rt <- 
  as_spectra(r.colourless, name_idx = 1, meta_idxs = 2:29)

r.white <- Points.raman %>% 
  filter(hue == "White")

spec.rw <- 
  as_spectra(r.white, name_idx = 1, meta_idxs = 2:29)

#Prepare a layout with 2 columns and 2 rows
par(mfrow=c(2,2))

#Plot the mean spectrum for each spectra group above with text annotations for the peaks in the first dark spectra
plot(mean(spec.rd), lwd = 1.2, col = "black", ylab = "Intensity", cex.lab = 1.5, cex.main = 1.5,
     xlab = "Wavenumber (cm^-1^)", main = "Average - Dark")
text(x=120, y=2900, labels="Quartz", font=2, cex = 1.5)
text(x=200, y=4300, labels="Quartz", font=2, cex = 1.5)
text(x=460, y=9250, labels="Quartz", font=2, cex = 1.5)
text(x=1285, y=6300, labels="D", font=2, cex = 2.0)
text(x=1600, y=4000, labels="G", font=2, cex = 2.0)

plot(mean(spec.rl), lwd = 1.2, col = "red", ylab = "Intensity", cex.lab = 1.5, cex.main = 1.5,
     xlab = "Wavenumber (cm^-1^)", main = "Average - Light")

plot(mean(spec.rt), lwd = 1.2, col = "purple", ylab = "Intensity", cex.lab = 1.5, cex.main = 1.5,
     xlab = "Wavenumber (cm^-1^)", main = "Average - Colourless")

plot(mean(spec.rw), lwd = 1.2, col = "blue", ylab = "Intensity", cex.lab = 1.5, cex.main = 1.5,
     xlab = "Wavenumber (cm^-1^)", main = "Average - White")
```

\linebreak

```{r raman-pca, fig.width=20, fig.height=15, fig.cap="PCA score plots for the raman spectra. The main explanatory power can be found in PC 1, which groups the data based on the quartz and D-G bands. As these are the main features in the data, the other components mainly describe the remaining noise, producing the arch-shape in the PCA. PC 4 does, however, also separate out the few samples that feature peaks at 700-800 cm^-1^ in the bottom left, that may result from Al, Si and OH vibrations."}
#perform PCA with SNV normalization and mean-center on the 
raman.pca <-
  prcomp(Points.raman[,c(30:321)], center = TRUE, scale = FALSE)

#prepare different basic score plots with fviz_pca_ind
basic_plot1 <-
  fviz_pca_ind(raman.pca, axes = c(1,2), label="none")

basic_plot2 <-
  fviz_pca_ind(raman.pca, axes = c(2,3), label="none")

#prepare labels for PCs
raman.pc1var <- round(summary(raman.pca)$importance[2,1]*100, digits=1)
raman.pc2var <- round(summary(raman.pca)$importance[2,2]*100, digits=1)
raman.pc4var <- round(summary(raman.pca)$importance[2,4]*100, digits=1)
raman.pc1lab <- paste0("PC1 (",as.character(raman.pc1var),"%)")
raman.pc2lab <- paste0("PC2 (",as.character(raman.pc2var),"%)")
raman.pc4lab <- paste0("PC3 (",as.character(raman.pc4var),"%)")

#RAMAN LOADING LINES
#par(mfrow=c(2,2))
#sapply(1:4, function(comp){
#  plot(row.names(raman.pca$rotation), raman.pca$rotation[,comp], type='l', lwd=2, 
#       main=paste("PC", comp), xlab="Wavelength (nm)", ylab="", cex.lab = 1.2, xaxp=c(92.88,1798.3,6))})

#bind the basic fviz plot for PC 1 and 2, and use as basis for a more customizeable plot in ggpplot
raman.p1 <- ggplot(cbind(basic_plot1$data, Points.raman[, c(8,10)]),
       aes(x=x, y=y, 
           shape = material, 
           fill = munsell_hue)) +
  geom_point(size=3) +
  theme_bw() +
  ggtitle("Raman") +
  labs(x = raman.pc1lab,
       y = raman.pc2lab) +
  scale_shape_manual(name = "Material",
                     values=pca.shapes) +
  scale_fill_manual(name = "Munsell hue",
                    breaks = c("Colourless", "2.5PB", "10B", "2.5GY", "5GY", "2.5Y", "5Y", "7.5Y", "10Y", "5YR", "N1", "N2", "N3", "N4", "N5", "N6", "N8", "N9"),
                    values=pca.colors) +
  guides(shape = guide_legend(override.aes = list(fill = "black"), 
                              ncol=2, 
                              title.position="top", 
                              title.hjust = 0.5),
         fill = guide_legend(override.aes = list(shape = 22,
                                                 fill = pca.colors,
                                                 color = "black",
                                                 size=3),
                             title.position="top", title.hjust = 0.5))+
  theme(legend.position = "none")

#bind the basic fviz plot for PC 1 and 3, and use as basis for a more customizeable plot in ggpplot
raman.p2 <- ggplot(cbind(basic_plot2$data, Points.raman[, c(8,10)]),
       aes(x=x, y=y, 
           shape = material, 
           fill = munsell_hue)) +
  geom_point(size=3) +
  theme_bw() +
  labs(x = raman.pc1lab,
       y = raman.pc4lab) +
  scale_shape_manual(name = "Material",
                     values=pca.shapes) +
  scale_fill_manual(name = "Munsell hue",
                    breaks = c("Colourless", "2.5PB", "10B", "2.5GY", "5GY", "2.5Y", "5Y", "7.5Y", "10Y", "5YR", "N1", "N2", "N3", "N4", "N5", "N6", "N8", "N9"),
                    values=pca.colors) +
  guides(shape = guide_legend(override.aes = list(fill = "black"), 
                              ncol=2, 
                              title.position="top", 
                              title.hjust = 0.5),
         fill = guide_legend(override.aes = list(shape = 22,
                                                 fill = pca.colors,
                                                 color = "black",
                                                 size=3),
                             title.position="top", title.hjust = 0.5))+
  theme(legend.position = "none")

comb <-
  raman.p1 + raman.p2 & theme(plot.title = element_text(size = 12, face = "bold", colour = "black"),
                              axis.title.x = element_text(size = 12, face = "bold", colour = "black"),
                              axis.title.y = element_text(size = 12, face = "bold", colour = "black"),
                              legend.title = element_text(size = 12, face = "bold", colour = "black"),
                              legend.background = element_rect(linetype = "solid", color = "black"),
                              legend.position = "bottom")

comb +
  plot_layout(nrow = 2, guides = "collect")

```

\linebreak

## XRF
Four samples had dimensions that were < 10 mm in size of their smallest dimension, and were therefore excluded from the analysis as there may be complications when comparing these to larger samples [@shackley11b pp.9-10]. The XRF spectrometer is not capable of detecting all of the common elements, and the software therefore calculates a balance (Bal) that represents the remaining unquantifiable elements in a sample. The elements represented in this unquantifiable group are all those with an atomic number lower than ^12^Mg (REF). Univariate analysis shows that SiO^2^ makes up the main bulk of the composition of the material, and two major groups can be distinguished based on this content (fig. \@ref(fig:xrf-violin-plot)). The majority of the readings resulted in a SiO^2^ content of less than 50%, and about ¼ of the readings (~ 27%) featured a SiO^2^ content closer to 90%. The “low” SiO^2^ group also feature a Bal between 40 – 70%.

The two first components of the PCA have an eigenvalue above 1 and the cumulative explained variance is ~ 81.3%. The loadings can be be seen in figure \@ref(fig:xrf-loadings). In the score plot of PC 1 and PC 2 two large clusters can be seen (fig. \@ref(fig:xrf-score-plot)). Interestingly, the smaller cluster on the left side in the score plot includes the samples with a SiO^2^ content higher than 50%, whilst the larger cluster includes those with lower SiO^2^ content. It does not seem to be ^14^Si that is responsible for this clustering, however, but rather ^26^Fe and ^40^Zr that are the main contributors. The “low ^14^Si” cluster shows a trend along PC 2 of quartzite with a darker hue grouping towards the bottom. This may indicate a possible relation to the ^26^Fe content of the material as ^26^Fe-bearing minerals such as hematite can contribute to the dark colour in quartzite [@howard05 p.711]. The brecciated quartz samples largely group together at the top of the “low ^14^Si” cluster, with a couple ending up closer towards the “high ^14^Si” cluster. The white (N9) material is fairly evenly distributed within the right cluster.

The group of measurements that are somewhat separated from the “low ^14^Si” cluster at the bottom are all but one from the rock painting site of *Finnforsberget*, which is located next to the Skellefte river. The *Finnforsberget* artefacts were sampled twice with object 491 having one measurement closer to the “low ^14^Si” group and the other the “high ^14^Si” group with ^13^Al estimated at 25 %. This artefact was sampled twice on each side, resulting in similar results both times.

\linebreak

```{r xrf-violin-plot, fig.width=15, fig.height=15, fig.fullwidth=TRUE, fig.cap="Violin plot of percentage element content in quartzite samples measured using pXRF. **n** represents no. of measurements above Limit of Detection (LOD) for that element out of a total of 474 measurements. Note the bimodal distribution of Si measurements, these groups are reflected in the score plot of the PCA"}
#Prepare data in long format for violin plots, exclude samples with a dimension smaller than 10 mm 
Points_long <-
  drop_na(Points.xrf %>%
    filter(max_length_mm >= 10 | max_width_mm >= 10) %>% 
    select(`12Mg`,`13Al`, `14Si`, `15P`, `16S`, `19K`, `20Ca`, `23V`, `25Mn`, `26Fe`, `30Zn`, `38Sr`, `39Y`, `40Zr`, `56Ba`, `81Ti`) %>%
    gather(key = "Element", value = "Value"))

# Add column with elemental groups for visualisation
Points_long <-
  Points_long %>%
    mutate(Group = case_when(
      endsWith(Element, "Mg") ~ "Lithophile",
      endsWith(Element, "Al") ~ "Lithophile",
      endsWith(Element, "Si") ~ "Lithophile",
      endsWith(Element, "P") ~ "Lithophile",
      endsWith(Element, "S") ~ "Chalcophile",
      endsWith(Element, "K") ~ "Lithophile",
      endsWith(Element, "Ca") ~ "Lithophile",
      endsWith(Element, "V") ~ "Lithophile",
      endsWith(Element, "Mn") ~ "Siderophile",
      endsWith(Element, "Fe") ~ "Siderophile",
      endsWith(Element, "Zn") ~ "Chalcophile",
      endsWith(Element, "Sr") ~ "Lithophile",
      endsWith(Element, "Y") ~ "Lithophile",
      endsWith(Element, "Zr") ~ "Lithophile",
      endsWith(Element, "Ba") ~ "Lithophile",
      endsWith(Element, "Ti") ~ "Chalcophile",
    ))

#Calculate and add value equal to 90% of the value range for each element, this is in preparation for 
#text annotations in final plot where a text will be positioned at this values height in each plot
Points_long <- 
  Points_long %>% 
    group_by(Element) %>% 
    dplyr::mutate(ypos = quantile(Value, 0.95))

#Prepare colors
loading.colors <- c("lightgoldenrod1", "darkorange", "lightpink1")

#Lookup for element labeller
#Ändra ordning så att de mer avvikande (Si, Mn etc) hamnar uppe till vänster
Elements <- c(
  '12Mg' = "Mg",
  '13Al' = "Al",
  '14Si' = "Si",
  '15P' = "P",
  '16S'= "S",
  '19K' = "K",
  '20Ca' ="Ca",
  '23V' = "V",
  '25Mn' = "Mn",
  '26Fe' = "Fe",
  '30Zn' = "Zn",
  '38Sr' = "Sr",
  '39Y' = "Y",
  '40Zr' = "Zr",
  '56Ba' = "Ba",
  '81Ti' = "Ti"
  )

#Prepare data frame with elements for annotating facets in final plot
na_values <- 
  data.frame("Element" = unique(Points_long$Element))

#Calculate and add number of missing values for each element             
na_values <- 
  na_values %>% 
    mutate(na_values = c(
      paste("n = ", length(which(!is.na(Points.xrf$`12Mg`)))),
      paste("n = ", length(which(!is.na(Points.xrf$`13Al`)))),
      paste("n = ", length(which(!is.na(Points.xrf$`14Si`)))),
      paste("n = ", length(which(!is.na(Points.xrf$`15P`)))),
      paste("n = ", length(which(!is.na(Points.xrf$`16S`)))),
      paste("n = ", length(which(!is.na(Points.xrf$`19K`)))),
      paste("n = ", length(which(!is.na(Points.xrf$`20Ca`)))),
      paste("n = ", length(which(!is.na(Points.xrf$`23V`)))),
      paste("n = ", length(which(!is.na(Points.xrf$`25Mn`)))),
      paste("n = ", length(which(!is.na(Points.xrf$`26Fe`)))),
      paste("n = ", length(which(!is.na(Points.xrf$`30Zn`)))),
      paste("n = ", length(which(!is.na(Points.xrf$`38Sr`)))),
      paste("n = ", length(which(!is.na(Points.xrf$`39Y`)))),
      paste("n = ", length(which(!is.na(Points.xrf$`40Zr`)))),
      paste("n = ", length(which(!is.na(Points.xrf$`56Ba`)))),
      paste("n = ", length(which(!is.na(Points.xrf$`81Ti`)))))
    )

#Merge na_values and Points_long data frames in order to get the y position that the text annotation should be displayed
#in for each element facet in final plot
na_values <- 
  merge(na_values, unique(Points_long[,c("Element", "ypos")]), by = "Element", all.x = TRUE)

#Prepare ylimits for each group in order to exclude extreme outliers in the violin plot graphic
ylims <- Points_long %>%
  group_by(Element) %>% 
  dplyr::summarise(y1 = 0, y2 = quantile(Value, 0.95))

#Prepare a list with formula for setting new ylimits for each group in ggh4x package "facetted_pos_scales() which makes it possible to work with facet_wrap()
scales <- c(
  Element == '12Mg' ~ scale_y_continuous(limits = c(as.vector(t(ylims[1,2:3])))),
  Element == '13Al' ~ scale_y_continuous(limits = c(as.vector(t(ylims[2,2:3])))),
  Element == '14Si' ~ scale_y_continuous(limits = c(as.vector(t(ylims[3,2:3])))),
  Element == '15P' ~ scale_y_continuous(limits = c(as.vector(t(ylims[4,2:3])))),
  Element == '16S' ~ scale_y_continuous(limits = c(as.vector(t(ylims[5,2:3])))),
  Element == '19K' ~ scale_y_continuous(limits = c(as.vector(t(ylims[6,2:3])))),
  Element == '20Ca' ~ scale_y_continuous(limits = c(as.vector(t(ylims[7,2:3])))),
  Element == '23V' ~ scale_y_continuous(limits = c(as.vector(t(ylims[8,2:3])))),
  Element == '25Mn' ~ scale_y_continuous(limits = c(as.vector(t(ylims[9,2:3])))),
  Element == '26Fe' ~ scale_y_continuous(limits = c(as.vector(t(ylims[10,2:3])))),
  Element == '30Zn' ~ scale_y_continuous(limits = c(as.vector(t(ylims[11,2:3])))),
  Element == '38Sr' ~ scale_y_continuous(limits = c(as.vector(t(ylims[12,2:3])))),
  Element == '39Y' ~ scale_y_continuous(limits = c(as.vector(t(ylims[13,2:3])))),
  Element == '40Zr' ~ scale_y_continuous(limits = c(as.vector(t(ylims[14,2:3])))),
  Element == '56Ba' ~ scale_y_continuous(limits = c(as.vector(t(ylims[15,2:3])))),
  Element == '81Ti' ~ scale_y_continuous(limits = c(as.vector(t(ylims[16,2:3]))))
)

#Violin plot of elemental content, trimmed, with fill based on elemental groups and number of values under LOD as text annotation
Points_long %>%
    ggplot(aes(x=Element, y=Value, fill = Group)) +
    geom_violin(trim = TRUE) +
    facet_wrap( ~ Element, scales = "free", labeller = labeller(Element = Elements)) +
    scale_fill_manual(values=loading.colors) +
    facetted_pos_scales(y = scales) +
    theme_bw() +
    geom_text(data=na_values, aes(x=0.2, y=ypos, label=na_values, fontface = "bold"), nudge_x = 0.4, inherit.aes = FALSE) +
    theme(
      legend.title = element_text(size = 12, face = "bold", colour = "black"),
      strip.text.x = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_blank(),
      plot.title = element_text(size=10, face = "bold", colour = "black", margin = margin(t = 10, b = -20), vjust=0.02, hjust=0.01),
      axis.title.y=element_blank(),
      axis.text.x=element_blank(),
      axis.ticks.x=element_blank())
```

\linebreak

```{r xrf-loadings, fig.width=8, fig.height=8, fig.cap="Loading vectors (P1-P2) with scores for each variable (element)"}
#filter on sample size (only include samples with smallest dimension >= 10mm), 
#select elements to include in pca, as well as columns to group by
xrf <-
  Points.xrf %>%
  filter(max_length_mm >= 10 & max_width_mm >= 10) %>% 
  select(reading_no, sample_id, munsell_hue, material, `12Mg`, `13Al`, `14Si`, `15P`, `16S`, `19K`, `20Ca`, `23V`, `25Mn`, `26Fe`, `30Zn`, `38Sr`, `39Y`, `40Zr`, `56Ba`, `81Ti`)

#log transform data
xrf.log <- log(xrf[,5:20])

#impute missing values
xrf.imp <- imputePCA(xrf.log, ncp = 2, scale = FALSE, method = "Regularized", maxiter = 1000)

#perform pca with mean centering and unit variance scaling
xrf.pca <- prcomp(xrf.imp$completeObs, center = TRUE, scale.=TRUE)

#Extract loadings from xrf.pca
loadings <- as.data.frame(xrf.pca$rotation[,1:2])
#As elements are stored as rownames, set them as first column so it can be used as a variable
loadings <- setDT(loadings, keep.rownames = "Element")[]
#Melt loadings from wide to long format
loadings.melted <- melt(loadings, id.vars="Element")
#Set names of columns
loadings <- setNames(loadings.melted, c("Element", "PC", "Value"))

#bar graph of the loadings (P1-P2) for the XRF data
loadings %>%
  ggplot(aes(x=fct_rev(Element), y=Value)) +
  geom_bar(
    stat = "identity", position = "identity",
    color = "black", fill = "darkgrey",
    width = 0.5) +
  facet_wrap( ~ PC, scales = "free") +
  theme_bw() +
  theme(plot.margin = margin(10,10,10,10),
        axis.title.y = element_blank(),
        axis.text.y = element_text(size=10, face="bold", colour = "black"),
        axis.title.x = element_blank(),
        strip.text.x = element_text(size=14, face="bold", colour="black")) +
  coord_flip()
```

\linebreak

```{r xrf-score-plot, fig.width=15, fullwidth=TRUE, fig.height=10, fig.cap="XRF score plot of PC 1 and PC 2. The smaller group to the left in the plot includes samples with higher Si content, while the larger group includes samples with Si content < 50%."}

#XRF score plot
ggbiplot(xrf.pca,
         choices=c(1,2),
         obs.scale = 1,
         var.scale = 1,
         group=NULL,
         varname.size = 4,
         labels.size=3,
         ellipse = FALSE,
         circle = FALSE) +
theme_minimal() +
geom_point(aes(shape=xrf$material, 
               fill = xrf$munsell_hue), 
           size = 3) +
coord_cartesian() +
scale_shape_manual(name = "Material",
                   values=pca.shapes) +
scale_fill_manual(name = "Munsell hue",
                  breaks = pca.munsell_hue,
                  values=pca.colors) +
guides(shape = guide_legend(override.aes = list(fill = "black"),
                            ncol=2,
                            title.position="top",
                            title.hjust = 0.5),
       fill = guide_legend(override.aes = list(shape = 22,
                                               fill = pca.colors,
                                               color = "black",
                                               size=3),
                           title.position="top", title.hjust = 0.5)) +
theme(plot.title = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 12, face = "bold", colour = "black"),
      legend.title = element_text(size = 12, face = "bold", colour = "black"),
      legend.background = element_rect(linetype = "solid", color = "black"),
      legend.position = "bottom")
```

\linebreak
<!--
```{r xrf-pca2, fig.width=15, fullwidth=TRUE, fig.height=10, fig.cap="XRF score plot of 1^st^ and 3rd components. The same groupings of low/high Si content remains along PC1, but PC3 seems to separate darker and light samples further in the low Si group."}
ggbiplot(xrf.pca,
         choices=c(1,3),
         obs.scale = 1,
         var.scale = 1,
         group=NULL,
         varname.size = 4,
         labels.size=3,
         ellipse = FALSE,
         circle = FALSE) +
theme_minimal() +
geom_point(aes(shape=xrf$material, 
               fill = xrf$munsell_hue), 
           size = 3) +
coord_cartesian() +
scale_shape_manual(name = "Material",
                   values=pca.shapes) +
scale_fill_manual(name = "Munsell hue",
                  breaks = pca.munsell_hue,
                  values=pca.colors) +
guides(shape = guide_legend(override.aes = list(fill = "black"),
                            ncol=2,
                            title.position="top",
                            title.hjust = 0.5),
       fill = guide_legend(override.aes = list(shape = 22,
                                               fill = pca.colors,
                                               color = "black",
                                               title.position="top", 
                                               title.hjust = 0.5))) +
theme(plot.title = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 12, face = "bold", colour = "black"),
      legend.title = element_text(size = 12, face = "bold", colour = "black"),
      legend.background = element_rect(linetype = "solid", color = "black"),
      legend.position = "bottom")
```
-->

# Discussion

\linebreak
 \linebreak

The most easily recognizable megascopic feature of petrographic material is its colour, a feature that has been one of the main means of grouping quartzite in North Swedish archaeological research [e.g. @forsberg85 p.60-64; @kaeck09 p. 53-55]. While it is true that some inference can be made based on colour, such as ^26^Fe-bearing minerals contributing to a darker colour, this is significantly more difficult with heterogeneous material. Should a quartzite where the matrix is mainly light gray in colour, but streaked with black mineral veins, be considered a *light* or *dark* quartzite? The results above reflects this complexity. While the above models tend to group samples of a similar hue together, especially in the case of NIRS, there are samples that break these trends and thus demonstrate the presence of structural and compositional features that do not necessarily align with the perceived colour of a material. Rather than assigning color based on how it is perceived visually it can be approached in a more systematic manner [e.g. @fairchild13]. Restricting the NIRS model to the visual range (~ 400 nm - 750 nm) opens up for some inference of color based on the absorbance of light within certain bands. However, as the data was collected using a contact probe the issue of how heterogeneous samples should be treated remains. This issue can be alleviated somewhat with a different sampling approach, for instance using a hyperspectral camera where different pixels can be selected and analysed separately [@sciuto19].

In the case of the Raman data the PCA is somewhat lacking in exploratory and descriptive power. Apart from the few samples that feature the 700 cm^-1^ - 800 cm^-1^ band peaks the major features in the data are all explained by the first component. As the intensity in Raman spectroscopy is not a direct quantitative indicator this makes it difficult to group materials based on the intensity of the D and G bands, making more of a binary separation of samples with or without *graphite peaks*. That is not to say that it is not possible to attempt quantification using Raman spectroscopy, but rather that this would necessitate a different experimental setup. That said, the presence of the *graphite peaks* may present an important piece of information in the identification of the potential geological origin of the material. Interestingly, mentions of *graphitic quartzite* can be found in an early 20th century description of an excursion to Storfjället, Tärna parish, by the Geological Society [@geologiska25], but otherwise does not seem to be a frequently used term in the modern day.

The material described as *brecciated quartz* remains relatively grouped together in all of the models, demonstrating some consistency with the original classification. As the samples group with other similarly semi-translucent material, however, it does not seem to distinguish itself enough in the models in order to consider it its own group. This is perhaps not so strange as the original classification by @ahman67 refers to the microscopic fragmented configuration of the quartz crystals as being the main distinguishing feature, and while it is possible to discern the presence of SiO structure through non-destructive spectroscopy, there have been no other discernible features related to this group at this stage. It is questionable whether all of the samples in this group truly exhibit brecciation, given the context under which they were collected and classified, which further complicates analysis of this material. Nevertheless, the results of the analysis suggests that there is some in-group variation, regardless of whether the material exhibits breccia.

\linebreak

# Conclusion

The results demonstrate clearly that non-destructive application of all three instrumentations (NIRS, Raman, XRF) are capable of petrographically characterizing some aspect of the quartz and quartzite, however, taken on their own they lack the resolution needed to provide a detailed enough fingerprint for classification. Nevertheless, they do provide a more in-depth understanding of the composition and structure of the material, as well as showcase the complexity in the petrographic character of quartz and quartzite materials. The study represents a good example of a multivariate exploratory approach to the characterization of petrographic material where there also is an absence of reference material. There is a need for methodologies that can be applied in regions with a complex geological setting and vague distribution of potential geological sources. Taking full advantage of the short integration time of these instrumentations in order to build large comprehensive datasets it is possible to discern spectral fingerprints using these methodologies.

\linebreak
 \linebreak

# Acknowledgements

I would like to thank my supervisors, Philip I. Buckland and Johan Linderholm, for valuable comments on previous versions of this paper.

\linebreak
 \linebreak

<!-- The following line inserts a page break  -->

\newpage

# References

\linebreak

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

<div id="refs"></div>

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```